---
title: "EDA"
author: "Amado"
date: '2022-09-22'
output: pdf_document
---

## **Load Libraries**

```{r setup, include=FALSE}
library(tidyverse)
library(gridExtra)
```

## **Load Dataset**

```{r cars}
df<- read.csv("C:\\Users\\PC\\Desktop\\Ciencias de datos\\Proyectos\\Car Evaluation\\data\\car_evaluation.csv")
```


## **Percentage Series**


```{r pressure, echo=FALSE}

serie <- function(...){
  
  serie <- df %>% group_by(...) %>% count() %>% ungroup() %>% mutate(percent=n/sum(n)*100)
  
  return(serie)
}

```


Develop a function with the objective of calculating the percentage of each category.



```{r}
attach(df)

serieBuying<- serie(buying)

serieMaint<-  serie(maint)

serieDoors<-   serie(doors)

seriePersons<-  serie(persons)

serieLugBoot<- serie(lug_boot)

serieSafety<-  serie(safety)

serieClass<-  serie(class)
```


## **Barplot**


```{r}
colors=c('#77dd77','#fdfd96','#84b6f4','#fdcae1')


barplot <- function(serie,x,...){
  
  serie %>% ggplot(aes(x=x,y=percent,fill=x)) + 
    geom_bar(stat = "identity",position = "dodge") +
    geom_text(aes(label=paste0(round(percent),"%"))) +
    theme_light() +
    scale_fill_manual(values = colors) +
    theme(legend.position = "none") +
    ...
  
}
```


```{r}

p1 <- barplot(serieBuying,serieBuying$buying,labs(x = "Buying",y = "Percent"))

p2 <- barplot(serieMaint,serieMaint$maint,labs(x = "Maint",y = "Percent"))

p3 <- barplot(serieDoors,serieDoors$doors,labs(x = "Doors",y = "Percent"))

p4 <- barplot(seriePersons,seriePersons$persons,labs(x = "Persons",y = "Percent"))

p5 <- barplot(serieLugBoot,serieLugBoot$lug_boot,labs(x = "Lug Boot",y = "Percent"))


p6 <- barplot(serieSafety,serieSafety$safety,labs(x = "Safety",y = "Percent"))

p7 <- barplot(serieClass,serieClass$class,labs(x = "Class",y = "Percent"))
```


## **Barplot Matrix**

```{r}
matrix<-grid.arrange(p1,p2,p3,p4,p5,p6,p7,top="Percentage by Category")
```


Most of the classes have a good proportion, except for the variable to be predicted, there is a clear disproportion between both classes. It can cause the classification model to have preference in the class with the most presence, for this reason it is convenient to make a balance adjustment.



```{r}

dir <- "C:\\Users\\PC\\Desktop\\Ciencias de datos\\Proyectos\\Car Evaluation\\img\\matrix_barplot.png"


ggsave(dir,matrix)
```

